Necessity is the mother of genetic invention

### 2017-11-03 ms.tex
threshold modifier

\documentclass[9pt, a4paper, twocolumn]{extarticle}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage[margin=1in]{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb,amsmath,amsthm}
\usepackage{longtable,booktabs}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\usepackage[auth-sc]{authblk}
\usepackage{titlesec}

@@

\date{\today} % Activate to display a given date or no date

@@

Another departure from the \emph{linear variation} assumption of the \emph{reduction principle} for mutation rates involves a mechanism commonly called \emph{stress-induced mutagenesis}, in which the mutation rate increases in individuals with poor fitness.
\citet{Ram2012} demonstrated that modifiers that increase the mutation rate of individuals with below-average fitness actually \textbf{increase} the population mean fitness, rather than decrease it.
Their analysis assumed an infinite population size and constant-viability selection, such that the only departure from the \emph{reduction principle} assumptions is due to the differential effect the modifier has on different genotypes.
\citet{Ram2012} stated that their result represents a departure from the \emph{reduction principle}, but did not explain this departure.
Their analysis was specific to a model that classified individuals by the number of mutant genes in their genotype, similar to those of~\citet{Kimura1966} and~\citet{Haigh1978}.

@@

\subsection*{Mutation modification model 1.}

@@

\subsection*{Mutation modification model 2.}

@@

\subsection*{Migration modification model.}

@@

\begin{mfp}
Let $\bar{\omega}^*$ be the leading eigenvalue of $(I-C+MC)D$, $u$ and $v$ be the corresponding left and right eigenvectors, and $[M]_k$ the $k$-th column of $M$. Then
\begin{equation}\label{eq:theorem}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k),
\end{equation}
or in simpler terms,
\begin{equation}
sign\frac{\partial \bar{\omega}^*}{\partial C_k} = 
sign(\bar{\omega}^* - D_k), 
\end{equation}
and therefore increased transition from type $k$ will increase the stable population mean fitness if and only if the fitness of type $k$ is below the stable population mean fitness.
\end{mfp}

@@

We turn our attention to a modifier that concurrently controls the transition out of a subset of types, $K$.
For example, \citet[Appendix~B]{Ram2012} considered individuals that are grouped by the number of their accumulated deleterious mutations, $k$, and a modifier that regulates the mutation rates in individuals with at least $\pi$ deleterious mutations.

Similarly, if the transition rates are controlled by a threshold rule 
\begin{equation}
C_k = \begin{cases}
\tau, & k \in K \\
1, & k \not\in K
\end{cases},
\end{equation}
then the sensitivity of the stable population mean fitness to changes in the
rate of transition out of types $K$ is
\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial \tau} = 
u\tr \Big( \sum_{k \in \mathbb{K{}}}{\frac{\partial (I - C + MC)D}{\partial \tau}} \Big) v = \\
\frac{1}{\tau} \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)}.
\end{aligned}
\end{equation}

\textbf{TODO} summarize result in words.

@@

\bibliographystyle{plainnat}
\bibliography{/Users/yoavram/Documents/library}

>>>>

\documentclass[9pt, a4paper, twocolumn]{extarticle}
\usepackage[margin=1in]{geometry} % See geometry.pdf to learn the layout options.
\geometry{a4paper}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx} % Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
% TeX will automatically convert eps --> pdf in pdflatex

\usepackage{amssymb,amsmath,amsthm}
\usepackage{longtable,booktabs}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\usepackage[auth-sc]{authblk}
\usepackage{titlesec}

@@

\date{\today}

@@

Another departure from the \emph{linear variation} assumption of the \emph{reduction principle} for mutation rates involves a mechanism commonly called \emph{stress-induced mutagenesis}, in which the mutation rate increases in individuals with poor fitness.
\citet{Ram2012} demonstrated that modifiers that increase the mutation rate of individuals with below-average fitness actually \textbf{increase} the population mean fitness, rather than decrease it.
Their analysis assumed an infinite population size and constant-viability selection, such that the only departure from the \emph{reduction principle} assumptions is due to the differential effect the modifier has on different genotypes.
\citet{Ram2012} stated that their result represents a departure from the \emph{reduction principle}, but did not explain this departure.
Their analysis was specific to a model that classified individuals by the number of mutant genes in their genotype, similar to those of~\citet{Kimura1966a} and~\citet{Haigh1978}.

@@

\subsection*{Mutation modification model 1.}\label{mut_model_1}

@@

\subsection*{Mutation modification model 2.}\label{mut_model_2}

@@

\subsection*{Migration modification model.}\label{mig_model_2}

@@

\begin{mfp}
Let $\bar{\omega}^*$ be the leading eigenvalue of $(I-C+MC)D$, $u$ and $v$ be the corresponding left and right eigenvectors, and $[M]_k$ the $k$-th column of $M$. Then
\begin{equation}\label{eq:theorem}
\frac{\partial \bar{\omega}^*}{\partial C_k} = 
\frac{u_k v_k}{C_k} (\bar{\omega}^* - D_k),
\end{equation}
or in simpler terms,
\begin{equation}\label{eq:sign_theorem}
sign\frac{\partial \bar{\omega}^*}{\partial C_k} = 
sign(\bar{\omega}^* - D_k), 
\end{equation}
and therefore increased transition from type $k$ will increase the stable population mean fitness if and only if the fitness of type $k$ is below the stable population mean fitness.
\end{mfp}

@@

% Threshold rule

We turn our attention to a modifier that concurrently controls the transition out of a subset of types, $K$.
For example, \citet[Appendix~B]{Ram2012} considered individuals that are grouped by the number of their accumulated deleterious mutations, $k$, and a modifier that regulates the mutation rates in individuals with at least $\pi$ deleterious mutations.
Similarly, if the transition rates are controlled by a threshold rule 
\begin{equation}
C_k = \begin{cases}
\tau, & k \in K \\
1, & k \not\in K
\end{cases},
\end{equation}
where $\tau > 0$, then the sensitivity of the stable population mean fitness to changes in the
rate of transition from types $K$ is
\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial \tau} = 
u\tr \Big( \sum_{k \in \mathbb{K{}}}{\frac{\partial (I - C + MC)D}{\partial \tau}} \Big) v = \\
\frac{1}{\tau} \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)},
\end{aligned}
\end{equation}
which can be simplified to
\begin{multline}
sign \frac{\partial \bar{\omega}^*}{\partial \tau} = 
sign \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)} = \\ 
sign \Big(\bar{\omega}^* - \frac{\sum_{k \in \mathbb{K}}{u_k v_k D_k}}{\sum_{k \in \mathbb{K}}{u_k v_k}}\Big), 
\end{multline}
a generalization of Eq.~\ref{eq:sign_theorem}.
Therefore, increased transition from types $K$ will increase the stable population mean fitness if and only if the average fitness of individuals descended from types $K$ is below the stable population mean fitness.

@@

\bibliographystyle{plainnat}
\bibliography{ms}
